---
title: "STAT433 HW1"
output: github_document
date: '2022-09-26'
author: "Xinyi Shi"
---

```{r}
library(dplyr)
library(nycflights13)
library(ggplot2)
```

1.How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
flights %>% filter(is.na(dep_time)) %>% summarise(nn=n())

flights %>% summary()
```
8255 flights have a missing dep_time, 8255 have a missing dep_delay, 8713 have a missing arr_time, 9430 have a missing arr_delay, 9430 have a missing air_time.
Those flights may failed to depart or being canceled, or they could be just lost data.


2.Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
flights %>% 
  mutate(new_dep_time=dep_time%%100+dep_time%/%100*60,
                   new_sched_dep_time=sched_dep_time%%100+sched_dep_time%/%100*60) %>% 
  summarise(new_dep_time,
            new_sched_dep_time)
```

3.Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()
```{r}
flights %>%
  mutate(date = lubridate::make_datetime(year, month, day)) %>%
  group_by(date) %>%
  summarise(cancelled = sum(is.na(dep_delay)), 
            n = n(),
            mean_dep_delay = mean(dep_delay,na.rm=TRUE),
             mean_arr_delay = mean(arr_delay,na.rm=TRUE)) %>%
  ggplot(aes(x= cancelled/n)) + 
  geom_point(aes(y=mean_dep_delay), colour='blue', alpha=0.5) + 
  geom_point(aes(y=mean_arr_delay), colour='red', alpha=0.5) +
  ylab('mean delay')
```

There is no strong evidence that there is some relationship. But generally there is a pattern that the higher cancelled/n is, the higher the mean delay is.